\texttt{escheR}: Unified multi-dimensional visualizations with Gestalt principles

### 2023-11-14 Manuscript/manuscript_BioAdv_format.tex
Clean up ref

The application of the proposed framework is not limited to \textit{in situ} visualizations of spatially-resolved transcriptomics data. It is broadly applicable to data mapped to any 2-dimensional coordinate system to simultaneously display multiple variables. Such systems include euclidean space (including spatial coordinate as a special case) and data-driven embedding space, for example, UMAP and $t$-SNE. To demonstrate, we applied the proposed visualization to address the challenge of simultaneously displaying cluster membership and gene expression in a single-cell UMAP plot. To address the overplotting problem, previous work proposed to apply hexagonal binning strategy to display the gene expression \citep{freytag_2020}. Here, the color-coded convex hulls are used to annotate different clusters of hexagons, where the cluster label of each hexagon is determined by the most popular label within each hexagon bin (\textbf{Figure~\ref{fig:embedding}A}). However, the convex hulls create substantial overlapping areas, creating confusion when interpreting cluster memberships of hexagons in the overlapping areas. To improve the interpretability of the visualization, we replace the convex hulls with color-coded hexagons boundaries (\textbf{Figure~\ref{fig:embedding}B}) to avoid possible membership confusion. We note that our contribution to improving the visualization is easily implemented without any modification of \texttt{schex} as both are built upon the Grammar of Graphics \citep{wilkinson_2012} standard.

>>>>

The application of the proposed framework is not limited to \textit{in situ} visualizations of spatially-resolved transcriptomics data. It is broadly applicable to data mapped to any 2-dimensional coordinate system to simultaneously display multiple variables. Such systems include euclidean space (including spatial coordinate as a special case) and data-driven embedding space, for example, UMAP and $t$-SNE. To demonstrate, we applied the proposed visualization to address the challenge of simultaneously displaying cluster membership and gene expression in a single-cell UMAP plot. To address the overplotting problem, previous work proposed to apply hexagonal binning strategy to display the gene expression \citep{freytag_2020}. Here, the color-coded convex hulls are used to annotate different clusters of \rr{hexagons, where the cluster label of each hexagon is determined by the most popular label within each hexagon bin} (\textbf{Figure~\ref{fig:embedding}A}). However, the convex hulls create substantial overlapping areas, creating confusion when interpreting cluster memberships of hexagons in the overlapping areas. To improve the interpretability of the visualization, we replace the convex hulls with color-coded hexagons boundaries (\textbf{Figure~\ref{fig:embedding}B}) to avoid possible membership confusion. We note that our contribution to improving the visualization is easily implemented without any modification of \texttt{schex} as both are built upon the Grammar of Graphics \citep{wilkinson_2012} standard.
